Constructing the transition graphs of interacting hysterons

### 2022-07-06 Paper_draft.tex
Removed out-of-place reference to number of graphs in sections on exchange symmetry/GoE states; minor wording change in line 123

To account for exchange symmetry, we fix the order in which hysterons 1, 2, ..., N flip up when increasing the external field from the ground state. For example, for $N=3$, we label the hysterons such that hysteron 3 flips up first from the state 000, hysteron 2 flips up first from 001, and trivially hysteron 1 flips up first from 011. Since there are $N!$ permutations of the order of $N$ up switching fields, accounting for exchange symmetry decreases the number of candidate graphs by a factor $N!$. Thus, for $N=2$, the number of candidate graphs is reduced by a factor 2 from 2916 to 1458.

@@

It often occurs that one or more of the possible states for $N$ hysterons cannot be reached from the ground state. As is customary\cite{juleslechenault2021}, we refer to states that cannot be reached as garden-of-Eden (GoE) states. We may choose to exclude GoE states in some cases, for instance in an experimental setup where one starts from the ground state. As a consequence of excluding GoE states, when two graphs only differ from one another by transitions that occur from GoE states, the graphs become duplicates. Therefore, when GoE states are ignored, the number of distinct graphs can turn out lower than the number of possible combinations of transitions. For $N=2$, ignoring GoE states  reduces the number of graphs from 1458 to 752. 

@@

In the case where hysterons are non-interacting, there are only $2N$ switching fields $\{u_i^{+, -}\}$. In addition, because we never compare an up switching field with a down switching fields in the absence of avalanches, the only orderings that can change the graph are the partial orderings of the up switching fields $\{u_i^+\}$ and of the down switching fields $\{u_i^-\}$. The number of orderings that can lead to a unique graph is then $N!\times N!$. When we fix the up switching fields to account for exchange symmetry, we find that the number of permutations is only $N!$.\\
Indeed there are $N!$ Preisach graphs for a collection of $N$ hysterons, each corresponding to a different ordering of the down switching fields\cite{terzimungan2020}. The ordering for each graph can be easily identified by considering the main loop: that is, the sequence of down transitions connecting the state with all $s_i = 1$ to that with all $s_i=0$. We accordingly use the main loop as an identifier for the Preisach graphs. For instance, for $N=2$ there are two Preisach graphs, corresponding to the main loops (1, 2) and (2, 1) \cite{terzimungan2020}.\\ 
We see that for the Preisach model, there is a one-to-one relationship between the permutations in the orderings of the switching fields $u_i^{+, -}$ and the possible transition graphs. When hysterons are allowed to interact, however, the situation changes. On one hand, there are then $2(2^N-1)$ states $U_i^{+, -}(S)$, and because of the possibility of avalanches, the ordering between all of the states must be considered. On the other hand, the state transition graph is only determined by a partial order between some of the switching fields. As a result, the number of permutations of the switching fields $(2(2^N-1))!$  for interacting hysterons greatly outnumbers the number of unique state transition graphs. For $N=2$, the number of permutations is $8! = 40320$ while we know the number of graphs to be only 13. For $N=3$, the number of permutations is as high as $24! \approx 6\times 10^{23}$.\\
We thus see that it is inefficient to find all graphs for an interacting system of hysterons by considering the permutations of all switching fields. Even so, we can make use of the fact that the state transition graph only depends on a partial ordering of the switching fields. Namely, we know that each graph has an underlying set of linear inequalities $U_i(S_A) > U_j(S_B)$, which we refer to as the design inequalities\cite{vhecke2021}. The solution space of the design inequalities provides the region in parameter space that leads to the given graph. In particular, if the design inequalities have no solution, then the given graph is not realizable.\\

>>>>

To account for exchange symmetry, we fix the order in which hysterons 1, 2, ..., N flip up when increasing the external field from the ground state. For example, for $N=3$, we label the hysterons such that hysteron 3 flips up first from the state 000, hysteron 2 flips up first from 001, and trivially hysteron 1 flips up first from 011. Since there are $N!$ permutations of the order of $N$ up switching fields, accounting for exchange symmetry decreases the number of candidate graphs by a factor $N!$.\\

@@

It often occurs that one or more of the possible states for $N$ hysterons cannot be reached from the ground state. As is customary\cite{juleslechenault2021}, we refer to states that cannot be reached as garden-of-Eden (GoE) states. We may choose to exclude GoE states in some cases, for instance in an experimental setup where one starts from the ground state. As a consequence of excluding GoE states, when two graphs only differ from one another by transitions that occur from GoE states, the graphs become duplicates. Therefore, when GoE states are ignored, the number of distinct graphs can turn out lower than the number of possible combinations of transitions.\\

@@

In the case where hysterons are non-interacting, there are only $2N$ switching fields $\{u_i^{+, -}\}$. In addition, because we never compare an up switching field with a down switching fields in the absence of avalanches, the only orderings that can change the graph are the partial orderings of the up switching fields $\{u_i^+\}$ and of the down switching fields $\{u_i^-\}$. The number of orderings that can lead to a unique graph is then $N!\times N!$. When we fix the up switching fields to account for exchange symmetry, we find that the number of permutations is only $N!$. Indeed there are $N!$ Preisach graphs for a collection of $N$ hysterons, each corresponding to a different ordering of the down switching fields\cite{terzimungan2020}.\\
 The ordering for each Preisach graph can be easily identified by considering the main loop: that is, the sequence of down transitions connecting the state with all $s_i = 1$ to that with all $s_i=0$. We accordingly use the main loop as an identifier for the Preisach graphs. For instance, for $N=2$ there are two Preisach graphs, corresponding to the main loops (1, 2) and (2, 1) \cite{terzimungan2020}.\\ 
We see that for the Preisach model, there is a one-to-one relationship between the permutations in the orderings of the switching fields $u_i^{+, -}$ and the possible transition graphs. When hysterons are allowed to interact, however, the situation changes. On one hand, there are then $2(2^N-1)$ states $U_i^{+, -}(S)$, and because of the possibility of avalanches, the ordering between all of the states must be considered. On the other hand, the state transition graph is only determined by a partial order between some of the switching fields. As a result, the number of permutations of the switching fields $(2(2^N-1))!$  for interacting hysterons greatly outnumbers the number of unique state transition graphs. For $N=2$, the number of permutations is $8! = 40320$ while we know the number of graphs to be only 13. For $N=3$, the number of permutations is as high as $24! \approx 6\times 10^{23}$. The gross overestimation of the number of graphs makes it a challenge to find and identify all graphs for an interacting system of hysterons.\\ 
Even so, we can make use of the fact that the state transition graph only depends on a partial ordering of the switching fields. Namely, we know that each graph has an underlying set of linear inequalities $U_i(S_A) > U_j(S_B)$, which we refer to as the design inequalities\cite{vhecke2021}. The solution space of the design inequalities provides the region in parameter space that leads to the given graph. In particular, if the design inequalities have no solution, then the given graph is not realizable.\\
